The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2016-12-06 manuscript/manuscript.Rnw
use short gene names in the intro

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all eutherian placentas, %
  %
  including genes with immune-modulating properties (\ac{ANXA2},
  \ac{ANXA1}, \ac{S100A11}, \ac{S100A10}, and \ac{LGALS1}), cell-cell
  interactions (\ac{LAMC1}, \ac{LUM}, and \ac{LGALS1}),
  invasion (\ac{GRB2} and \ac{RALB}) and syncytialization
  (\ac{ANXA5} and \ac{ANXA1}). %
  We also identified multiple pre-eclampsia associated genes which are
  differentially expressed in \textit{Homo sapiens} when compared to
  the other %
  \Sexpr{star.logs[,length(unique(species))-1]} species. %
  %
  Multiple genes are significantly associated with placenta
  morphology, including \ac{EREG} and \ac{WNT5A} which are both
  associated with placental shape.

@@

Multiple components of annexin complexes are included in the core
transcriptome, including \ac{ANXA2}, \ac{ANXA1},
\ac{S100A11}, \ac{S100A10}. ANXA1 is involved in the
resolution of the inflammation response through inhibition of
phospholipase A2 activity and signaling through the formyl peptide
receptor family, and has previously been hypothesized to be important
for the maintenance of the anti-inflammatory state during pregnancy
\cite{Hutchinson.ea2011:Molecularregulatorsofresolution}. ANXA2 is
known to interact with S100A10, S100A6, S100A11, and S100A4
\cite{Liu.ea2015:AnnexinA2complexeswith}, and there is evidence for
its interaction with S100P as well; it is involved in cell-cell
interaction as well as vesicle trafficking and von Willibrand Factor
secretion. The expression regulation of another core gene,
\ac{TIMP3}, has been implicated in pre-eclampsia
\cite{Ma.ea2014:Down-regulationofTIMP3leads,Xiang.ea2013:PromoterhypomethylationofTIMP3}.

>>>>

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all eutherian placentas, %
  %
  including genes with immune-modulating properties (\acs{ANXA2},
  \acs{ANXA1}, \acs{S100A11}, \acs{S100A10}, and \acs{LGALS1}), cell-cell
  interactions (\acs{LAMC1}, \acs{LUM}, and \acs{LGALS1}),
  invasion (\acs{GRB2} and \acs{RALB}) and syncytialization
  (\acs{ANXA5} and \acs{ANXA1}). %
  We also identified multiple pre-eclampsia associated genes which are
  differentially expressed in \textit{Homo sapiens} when compared to
  the other %
  \Sexpr{star.logs[,length(unique(species))-1]} species. %
  %
  Multiple genes are significantly associated with placenta
  morphology, including \acs{EREG} and \acs{WNT5A} which are both
  associated with placental shape.

@@

Multiple components of annexin complexes are included in the core
transcriptome, including \ac{ANXA2}, \ac{ANXA1},
\ac{S100A11}, \ac{S100A10}. ANXA1 is involved in the
resolution of the inflammation response through inhibition of
phospholipase A2 activity and signaling through the formyl peptide
receptor family, and has previously been hypothesized to be important
for the maintenance of the anti-inflammatory state during pregnancy
\cite{Hutchinson.ea2011:Molecularregulatorsofresolution}. ANXA2 is
known to interact with S100A10, S100A6, S100A11, and S100A4
\cite{Liu.ea2015:AnnexinA2complexeswith}, and there is evidence for
its interaction with S100P as well; it is involved in cell-cell
interaction as well as vesicle trafficking and von Willibrand Factor
secretion. The expression regulation of another core gene,
\ac{TIMP3}, has been implicated in pre-eclampsia
\cite{Ma.ea2014:Down-regulationofTIMP3leads}.
